---
title: "Tools for Quantitative Archaeology – in R"
authors:
  - "Joe Roe <<joe@joeroe.io>>"
date: "`r Sys.Date()`"
output: github_document
---

<!-- Generated from README.Rmd -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "readme_figures/"
)

library("tidyverse", warn.conflicts = FALSE)
library("gt")
```

[Tools for Quantitative Archaeology](http://tfqa.com) (TFQA) is a collection of DOS programs developed by Keith Kintigh to perform statistical analyses used in archaeology.
TFQA includes 50 programs and so gives a good representation of the range of analyses used by archaeologists.
The purpose of this document is to track which of these analyses are currently available in R packages.
We hope it will be useful in both porting TFQA-based analyses to R, and in highlighting which methods are not yet implemented in R packages.

The table below presents a list of [TFQA programs](http://tfqa.com/programs.htm) and their equivalent functions in R.
By "equivalent", we mean R functions that provide substantially the same functionality as the original TFQA program with a similar high-level user interface; 
we can assume that all of the analyses listed can be performed in R if the user is prepared to reimplement them themselves. 
The list of R equivalents is also not intended to be exhaustive.
The packages/functions listed are a subjective assessment of the "best" (most complete/widely used/actively maintained) way to get the same results as the TFQA program using R.
See [open-archeo.info](http://open-archaeo.info/) for a general list of archaeology-related R packages.

TFQA is still available and actively maintained.
Some of the more recently added programs are also available as free and open source software (see <https://github.com/kintigh>).
Matt Peeples has also [ported many to R scripts](https://mattpeeples.net/?page_id=656).
This list should therefore be seen as list of alternatives to TFQA (for those who prefer R, open source software, and/or cannot run DOS programs), rather than superseding it.

This is a work-in-progress.
Please feel free to contribute by [editing the table](/sslarch/tfqar/blob/main/tfqar.csv) and submitting a pull request, or [opening an issue](/sslarch/tfqar/issues) with suggestions.

## R equivalents of TFQA programs

Generated from [tfqar.csv](/sslarch/tfqar/blob/main/tfqar.csv).

```{r tfqar-functions, include=FALSE}
md_collapse <- function(x, nwords = 10) {
  summary <- word(x, end = nwords)
  summary[is.na(summary)] <- x[is.na(summary)]
  collapsed <- paste0("<details>",
                      paste0("<summary>", summary, "...</summary>"),
                      x,
                      "</details>",
                      sep = "\n")
  x[x != summary] <- collapsed[x != summary]
  return(x)
}

plain_list <- function(x) {
  x <- str_remove_all(x, " ")
  x <- str_replace_all(x, ",", "<br />")
  return(x)
}
```

```{r tfqar-table, echo=FALSE}
tfqar <- read_csv("tfqar.csv", col_types = "ccccccc")

tfqar %>% 
  group_by(tfqa_group) %>% 
  gt() %>% 
  cols_label(tfqa_program = "TFQA Program",
             tfqa_description = "Description",
             r = "Available in R?",
             r_package = "R package(s)",
             r_function = "R function(s)",
             notes = "Notes") %>% 
  sub_missing(everything(), missing_text = "") %>% 
  text_transform(md_collapse, cells_body(tfqa_description)) %>% 
  text_transform(plain_list, cells_body(r_function)) %>% 
  fmt_markdown(c(r_package, notes)) %>% 
  tab_source_note(md("TFQA program descriptions copied from <http://tfqa.com/programs.htm>")) %>% 
  tab_style(cell_text(weight = "bold"), 
            list(cells_column_labels(everything()), cells_row_groups())) %>%
  tab_style(cell_text(v_align = "top"), cells_body()) %>%
  as_raw_html()
```